---
title: "[Your title (e.g Exercise 4.3.2)]"
author:
  - name: [First author name]
    #url: https://example.com/norajones
    affiliation: cand.merc (OSCA)
    #affiliation_url: https://example.com/spacelysprokets
  - name: [Second author name]
    affiliation: cand.merc (?)
  - name: [Third author name]
    affiliation: cand.merc (?)
output: 
  html_document:
    toc: true
    toc_depth: 2
date: "`r Sys.Date()`"
---


```{r setup, include=FALSE}
# set default chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  autodep = TRUE
  )
```

<!-- Some css for the comments (delete) -->
```{css, echo=FALSE}
.comment {
  color: gold;
  border-style: dashed;
  border-width: thin;
  padding: 10px;
}
```


<div class="comment">This is a template for a TFA exercise. Modify it as needed. Text in gold are comments. Some example text, comments and code are given. Delete as needed.</div>

## Introduction

We load the following packages:

```{r}
library(tidyverse)
```

<div class="comment">
Maybe a short intro about the exercise and how data are imported.

Some help links:

- To use this template you must have installed the tfa package (`remotes::install_github("bss-osca/tfa-package")`). Go to **File > New File > R Markdown...** and select the **From template** in the left column and then **TFA Exercise**.
- For Markdown basics see **Help > Markdown Quick Reference** in RStudio.
- For R Markdown basics see the **Help > Cheatsheets > R Markdown Cheat Sheet** in RStudio.
- Learn more about he Distill format for R Markdown at <https://rstudio.github.io/distill>.
- To see the possible options for R chunks see <https://yihui.org/knitr/options/>.
</div>

Consider a problem of determining the best sequencing of jobs on a machine. A set of startup costs are given for 5 machines:

```{r}
startup_costs <- c(27, 28, 32, 35, 26)
startup_costs
```

Moreover, when changing from one job to another job the setup costs are given as:

```{r}
setup_costs <- matrix(c(
  NA, 35, 22, 44, 12,
  49, NA, 46, 38, 17,
  46, 12, NA, 29, 41,
  23, 37, 31, NA, 26,
  17, 23, 28, 34, NA), 
  byrow = T, nrow = 5)
setup_costs
```

The goal of the problem is to determine a sequence of jobs which minimizes the total setup cost including the startup cost.

One possible way to find a sequence is the use a greedy strategy:

```
Greedy Algorithm
Step 0: Start with the job which has minimal startup cost.
Step 1: Select the next job as the job not already done 
        with minimal setup cost given current job. 
Step 2: Set next job in Step 1 to current job and 
        go to Step 1 if not all jobs are done.
```

The greedy algorithm can be implemented as:

```{r}
greedy <- function(startup, setup) {
  jobs <- nrow(setup)
  cur_job <- which.min(startup)
  cost <- startup[cur_job]
  cat("Start job:", cur_job, "\n")
  job_seq <- cur_job
  setup[, cur_job] <- NA
  # browser()
  for (i in 1:(jobs-1)) {
    next_job <- which.min(setup[cur_job, ])
    cat("Next job:", next_job, "\n") 
    cost <- cost + setup[cur_job, next_job]
    job_seq <- c(job_seq, next_job)
    cur_job <- next_job
    setup[, cur_job] <- NA
  }
  # print(setup)
  return(list(seq = job_seq, cost = cost))
}
greedy(startup_costs, setup_costs)
```

First, the job with minimum startup cost is found using function `which.min` and we define cost a the startup cost. We use `cat` to make some debugging statements and initialize `job_seq` with the first job. Next we have to have a way ignoring jobs already done. We do that here by setting the columns of setup cost equal to NA for jobs already done. Hence they will not be selected by `which.min`. The `for` loop runs 4 times and selects jobs and accumulate the total cost.


## Question 1

A well-known better strategy is to:

```
Better Algorithm
Step 0: Subtract minimum of startup and setup cost for each job from setup and 
        startup costs (that is columnwise)
Step 1: Call the greedy algorithm with the modified costs. Note that the total 
        cost returned has to be modified a bit.
```

*Implement a `better` function calculating a better strategy. Hint: to find the minimum column costs you may use `apply(rbind(startup, setup), 2, min, na.rm = T)`.*

The function becomes:

```{r}
better <- function(startup, setup) {
  jobs <- nrow(setup)
  min_col_val <- apply(rbind(startup, setup), 2, min, na.rm = T)
  startup <- startup - min_col_val
  min_mat <- matrix(rep(min_col_val, jobs), 
                    ncol = jobs, byrow = T)
  setup <- setup - min_mat
  lst <- greedy(startup, setup)
  lst$cost <- lst$cost + sum(min_col_val)
  return(lst)
}
```

If we run the function the job sequence becomes:

```{r}
res <- better(startup_costs, setup_costs)
```

That is we use job sequence `r res$seq` with a cost of `r res$cost`.


## Question 2

*Assume that the startup costs are 27, 28, 55, 120 and 26 instead. How will that affect the job sequence using the better strategy?*

We modify the startup costs and run `better`:

```{r}
startup_costs <- c(27, 28, 55, 120, 26)
res <- better(startup_costs, setup_costs)
res
```

The job sequence now becomes `r res$seq` with a cost of `r res$cost`.


## Question 3

*Make a plot of the job sequence*

```{r, layout="l-body-outset", fig.asp = 0.5, eval=TRUE}
library(ggraph)
library(tidygraph)
edges <- tibble(from = res$seq[1:4], to = res$seq[2:5])
graph <- as_tbl_graph(edges) 
ggraph(graph, layout = 'linear') + 
  geom_edge_link(arrow = arrow(length = unit(4, 'mm')), 
                 end_cap = circle(4, 'mm'), 
                 start_cap  = circle(4, 'mm')) + 
  geom_node_point(colour = "gray", size = 9, alpha = 0.5) +
  geom_node_text(aes(label=name)) +
  theme_graph()
```


## Question 4

*What is the sum of all setup costs?*

```{r}
scost <- sum(setup_costs, na.rm=TRUE)


```

The sum of all setup costs er `r scost`


```{r}

# remotes::install_github("bss-osca/tfa-package", build = FALSE)  # run if tfa not installed
library(tidyverse)
dat <- read_csv(system.file("extdata/titanic.csv", package = "tfa"))

Data <- (mutate(dat, male=ifelse(sex=="male", TRUE, FALSE)))
Data


library(skimr)
skim(Data)
#Ved at skimme data kan man få et hurtigt overblik over det. Den laver summeringer af alle variablerne og beregner let statistik

#Vi kan løse det her ved at lave grupperinger

Data %>% 
  group_by(sex, survived) %>% 
  count() %>% 
  group_by(sex) %>% 
  mutate(rate=n/sum(n))



#tæl hvor mange børn der er under 19 år
Data

Data %>% 
  group_by(survived, age < 19) %>% 
  count() %>% 
  group_by(survived) %>% 
  mutate(rate=n/sum(n))

##Q4 Alternativ løsning
data %>%
  count(age < 19)
#196
  filter(age < 19 , survived == TRUE) %>%
  count()
# 95


#Hvor mange overlevede på 1st class
Data %>% 
  group_by(pclass, survived) %>% 
  count() %>% 
  group_by(pclass) %>% 
  mutate(rate=n/sum(n))


#Hvor mange overlevede i life boat
Data

Data %>% 
  filter(!is.na(boat)) %>% 
  summarise(rate=1-sum(survived)/n())
  print()
  
  
#How many persons with Hansen in their name survived?
  Data %>% 
  filter(str_detect(name, "Hansen"))
    
    
  
```

```{r}
library(skimr)
library(tidyverse)
path <- system.file("extdata/profit_raw.csv", package = "tfa")
profit <- read_csv(path)
skim(profit)
profit %>% distinct(Quarter)

profit <- profit %>%  
  mutate(Quarter = str_remove(Quarter, "q") %>% str_remove("Q") %>% as.numeric())
profit %>% distinct(Quarter)



profit

profit <- profit %>% 
  mutate(Revenue = str_replace_all(Revenue, c("unknown" = "?", "Unknown" = "?"))) %>% 
  mutate(Revenue = na_if(Revenue, "?")) %>% 
  mutate(Revenue = str_remove(Revenue, fixed("$ ")) %>% as.numeric())

profit <- profit %>% 
  mutate(`Product Cost` = str_replace_all(`Product Cost`, c("unknown" = "?", "Unknown" = "?"))) %>% 
  mutate(`Product Cost` = na_if(`Product Cost`, "?")) %>% 
  mutate(`Product Cost` = str_remove(`Product Cost`, fixed("$ ")) %>% as.numeric())

profit <- profit %>% 
  mutate(`Customer Service Cost` = str_replace_all(`Customer Service Cost`, c("unknown" = "?", "Unknown" = "?"))) %>% 
  mutate(`Customer Service Cost` = na_if(`Customer Service Cost`, "?")) %>% 
  mutate(`Customer Service Cost` = str_remove(`Customer Service Cost`, fixed("$ ")) %>% as.numeric())

profit <- profit %>% 
  mutate(Profit = str_replace_all(Profit, c("unknown" = "?", "Unknown" = "?"))) %>% 
  mutate(Profit = na_if(Profit, "?")) %>% 
  mutate(Profit = str_remove(Profit, fixed("$ ")) %>% as.numeric())

profit %>% distinct(Profit)
profit

#Make use of across function, str_remove and mutate

profit <- profit %>% 
  mutate(Quarter = str_remove(Quarter, "q") %>% str_remove("Q") %>% as.numeric()) %>% 
  mutate(across(Revenue:Profit, ~str_replace_all(.x,c("unknown" = "?", "Unknown" = "?")))) %>%  #Her bruges .x for at vælge hver element i across funktionen. Altså, .x = Revenue og dernæst... ... Profit.
  mutate(across(Revenue:Profit, ~na_if(.x, "?"))) %>%
  mutate(across(Revenue:Profit, ~str_remove(.x, fixed("$ ")) %>% as.numeric()))


profit


#Validate the calculation for profit. If the difference is between 0 and 1 then recalculate the Profit column. Also ignore any na.
profit <- profit %>% 
  mutate(Profit_calc = Revenue - `Product Cost` - `Customer Service Cost`, Diff = abs(Profit - Profit_calc)) %>%  
  mutate(Profit = if_else(Diff > 0 & Diff <= 1, Profit_calc, Profit, Profit) ) 
profit %>% filter(Diff == 1, is.na(Profit_calc))  # check

profit



#Find NA missing values henover flere kolonner og for hver række
profit <- profit %>% 
  rowwise() %>% 
  mutate(ct_na = sum(is.na(c_across(Revenue:Profit)))) %>% 
  ungroup()
profit %>% filter(ct_na >= 1)


#Recalculate Revenue and Profit if there are one NA value. If there are multiple then remove
profit <- profit %>% 
  rowwise() %>% 
  mutate(c_na = sum(is.na(c_across(Revenue:Profit)))) %>% 
  mutate(Revenue = if_else(is.na(Revenue) & c_na == 1, Profit + `Product Cost` + `Customer Service Cost`, Revenue, Revenue),
         `Product Cost` = if_else(is.na(`Product Cost`) & c_na == 1, - Profit + Revenue - `Customer Service Cost`, `Product Cost`), 
         `Customer Service Cost` = if_else(is.na(`Customer Service Cost`) & c_na == 1, - Profit + Revenue - `Product Cost`, `Customer Service Cost`)) %>% 
  select(Quarter:Profit)


profit

#Check if the calculated profit is equal to the actual profit. If yes then there should be no difference.
profit %>% 
  mutate(Profit_calc = Revenue - `Product Cost` - `Customer Service Cost`, Diff = abs(Profit - Profit_calc)) %>%  
  filter(Diff > 0)

profit

#Find NA values

profit %>% 
  rowwise() %>% 
  mutate(c_na = sum(is.na(c_across(Revenue:Profit)))) %>% 
  filter(c_na > 0)

#Find the two highest profit rows for each quarter.
profit %>% 
  group_by(Quarter) %>% 
  slice_max(Profit,n=2)


#Find the two best customers with highest profit in each quarter
profit %>% 
  group_by(`Customer ID`) %>% 
  slice_max(Profit, n=2) %>% 
  arrange(`Customer ID`)


#Find the product line, customer, channel, country and quarter with the highest profit.
profit %>% 
  group_by(`Product Line`) %>% 
  summarise(Profit = sum(Profit)) %>% 
  slice_max(Profit)

summarise_profit  <- function(data, group_var, summarise_var) {
  data %>%
    group_by(across({{ group_var }})) %>% 
    summarise(across({{ summarise_var }}, sum)) %>% 
    slice_max(Profit)
}
summarise_profit(profit, `Product Line`, Profit)
#> # A tibble: 1 × 2
#>   `Product Line`  Profit
#>   <chr>            <dbl>
#> 1 Credit Products  31022
summarise_profit(profit, `Customer ID`, Profit)
#> # A tibble: 1 × 2
#>   `Customer ID` Profit
#>   <chr>          <dbl>
#> 1 WEM            18942
# ... repeat

# Using purrr package to get a single tibble (if interested in the purrr package)
val <- names(profit)[1:5]
max_profit <- map_df(
  val, 
  ~{
      tmp <- profit %>% 
        group_by(.data[[.x]]) %>% 
        summarise(Profit = sum(Profit), .groups = "drop") %>% 
        slice_max(Profit)
      tibble(by = .x, 
        best = as.character(tmp[[1,1]]),
        profit = tmp[[1,2]] 
      )
    } 
  )
max_profit 
  
  


#Same customer ID in multiple countries?
profit %>% 
  group_by(`Customer ID`) %>% 
  distinct(Country) %>% 
  count(`Customer ID`)
  

#Sort the data decreasing with respect to profit and next revenue.
profit
profit %>% 
  arrange(desc(Profit),Revenue)



#Which product line has the highest and lowest total cost
profit %>% 
  group_by(`Product Line`) %>% 
  summarise(cost = sum(`Product Cost` + `Customer Service Cost`)) %>% 
  # print() %>%   # if want a peek before slicing 
  slice_max(cost)

profit %>% 
  group_by(`Product Line`) %>% 
  summarise(cost = sum(`Product Cost`)) %>% 
  slice_min(cost)


#Assume that customer service cost increases with 5%. How will that affect the profit for each product line?
profit %>%
  mutate(cust_cost_new = `Customer Service Cost` * 1.05, 
         profit_new = Revenue - cust_cost_new - `Product Cost`) %>% 
  group_by(`Product Line`) %>% 
  summarise(cust_cost = sum(`Customer Service Cost`), profit = sum(Profit), 
            cust_cost_new = sum(cust_cost_new), profit_new = sum(profit_new),
            profit_decrease = profit_new - profit)
```


```{r}
library(tidyverse)
fisheries <- read_csv(system.file("extdata/fisheries.csv", package = "tfa"))
fisheries

continents <- read_csv(system.file("extdata/continents.csv", package = "tfa"))
continents

fisheries %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))

#Use a mutating join to add a continent column to the fisheries dataset. Are there some countries which do not belong to a continent? vi bruger her left join for at bibeholde alle kolonner i fisheries så vi kun kan komme til at tilføje kolonner (continents)
fisheries
fisheries <- fisheries %>% 
  left_join(continents)

fisheries
#Find countries which don't belong to a continent
fisheries %>% 
  anti_join(continents)

fisheriesfilter <- fisheries %>% 
  filter(total > 100000)
fisheriesfilter

#Se hvilke lande som ikke er i et kontinent efter filtrering
fisheriesfilter %>% filter(is.na(continent))


#Fiks de lande som ikke er i et kontinent. Dvs. tilføj dem til det tættest kontinent
fisheriesfilter %>% filter(is.na(continent))
fisheriesfilter <- fisheriesfilter %>%
  mutate(continent = case_when(
    country == "Democratic Republic of the Congo" ~ "Africa",
    country == "Hong Kong"                        ~ "Asia",
    country == "Myanmar"                          ~ "Asia",
    TRUE                                          ~ continent
    )
  )
fisheriesfilter %>% filter(is.na(continent))


#Add column:
fisheriesfilter <- fisheriesfilter %>% 
  mutate(aquaculture_perc = aquaculture / total)
#Explanation: Andelen som aquaculture udgører af den total produktion virksomheden har haft.


#Beregn mean for hvert continent af aquaculture perc. Såvel som min og max.
fisheriesfilter %>%                 
  group_by(continent) %>%                      
  summarize(mean_ap = mean(aquaculture_perc),min=min(aquaculture_perc),max=max(aquaculture_perc)) 





```

## Colophon 
<!-- Always keep this section for reproducibility -->

This report has been created inside [RStudio](http://www.rstudio.com/ide/) using [R Markdown](https://rmarkdown.rstudio.com/). 

The report was built using:

```{r message = FALSE, warning = FALSE, echo = FALSE}
session <- devtools::session_info()
session$platform
```

Along with these packages:

```{r message = FALSE, warning = FALSE, echo = FALSE}
session$packages %>% 
  select(package, loadedversion, date, source) %>% 
  DT::datatable(rownames = FALSE,
                class = 'cell-border stripe',
                filter = list(position = 'top'),
                options = list(pageLength = 5, 
                           autoWidth = FALSE,
                           bInfo = FALSE,
                           paging = TRUE))
```








