---
title: "[Your title (e.g Exercise 4.3.2)]"
author:
  - name: [First author name]
    #url: https://example.com/norajones
    affiliation: cand.merc (OSCM)
    #affiliation_url: https://example.com/spacelysprokets
  - name: [Second author name]
    affiliation: cand.merc (?)
  - name: [Third author name]
    affiliation: cand.merc (?)
output: 
  html_document:
    toc: true
    toc_depth: 2
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
# set default chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  autodep = TRUE
  )
```

<!-- Some css for the comments (delete) -->
```{css, echo=FALSE}
.comment {
  color: gold;
  border-style: dashed;
  border-width: thin;
  padding: 10px;
}
```


<div class="comment">This is a template for a TFA exercise. Modify it as needed. Text in gold are comments. Some example text, comments and code are given. Delete as needed.</div>

## Introduction

We load the following packages:

```{r}
library(tidyverse)
```

<div class="comment">
Maybe a short intro about the exercise and how data are imported.

Some help links:

- To use this template you must have installed the tfa package (`remotes::install_github("bss-osca/tfa-package")`). Go to **File > New File > R Markdown...** and select the **From template** in the left column and then **TFA Exercise**.
- For Markdown basics see **Help > Markdown Quick Reference** in RStudio.
- For R Markdown basics see the **Help > Cheatsheets > R Markdown Cheat Sheet** in RStudio.
- Learn more about he Distill format for R Markdown at <https://rstudio.github.io/distill>.
- To see the possible options for R chunks see <https://yihui.org/knitr/options/>.
</div>

Consider a problem of determining the best sequencing of jobs on a machine. A set of startup costs are given for 5 machines:

```{r}
startup_costs <- c(27, 28, 32, 35, 26)
startup_costs
```

Moreover, when changing from one job to another job the setup costs are given as:

```{r}
setup_costs <- matrix(c(
  NA, 35, 22, 44, 12,
  49, NA, 46, 38, 17,
  46, 12, NA, 29, 41,
  23, 37, 31, NA, 26,
  17, 23, 28, 34, NA), 
  byrow = T, nrow = 5)
setup_costs
```

The goal of the problem is to determine a sequence of jobs which minimizes the total setup cost including the startup cost.

One possible way to find a sequence is the use a greedy strategy:

```
Greedy Algorithm
Step 0: Start with the job which has minimal startup cost.
Step 1: Select the next job as the job not already done 
        with minimal setup cost given current job. 
Step 2: Set next job in Step 1 to current job and 
        go to Step 1 if not all jobs are done.
```

The greedy algorithm can be implemented as:

```{r}
greedy <- function(startup, setup) {
  jobs <- nrow(setup)
  cur_job <- which.min(startup)
  cost <- startup[cur_job]
  cat("Start job:", cur_job, "\n")
  job_seq <- cur_job
  setup[, cur_job] <- NA
  # browser()
  for (i in 1:(jobs-1)) {
    next_job <- which.min(setup[cur_job, ])
    cat("Next job:", next_job, "\n") 
    cost <- cost + setup[cur_job, next_job]
    job_seq <- c(job_seq, next_job)
    cur_job <- next_job
    setup[, cur_job] <- NA
  }
  # print(setup)
  return(list(seq = job_seq, cost = cost))
}
greedy(startup_costs, setup_costs)
```

First, the job with minimum startup cost is found using function `which.min` and we define cost a the startup cost. We use `cat` to make some debugging statements and initialize `job_seq` with the first job. Next we have to have a way ignoring jobs already done. We do that here by setting the columns of setup cost equal to NA for jobs already done. Hence they will not be selected by `which.min`. The `for` loop runs 4 times and selects jobs and accumulate the total cost.


## Question 1

A well-known better strategy is to:

```
Better Algorithm
Step 0: Subtract minimum of startup and setup cost for each job from setup and 
        startup costs (that is columnwise)
Step 1: Call the greedy algorithm with the modified costs. Note that the total 
        cost returned has to be modified a bit.
```

*Implement a `better` function calculating a better strategy. Hint: to find the minimum column costs you may use `apply(rbind(startup, setup), 2, min, na.rm = T)`.*

The function becomes:

```{r}
better <- function(startup, setup) {
  jobs <- nrow(setup)
  min_col_val <- apply(rbind(startup, setup), 2, min, na.rm = T)
  startup <- startup - min_col_val
  min_mat <- matrix(rep(min_col_val, jobs), 
                    ncol = jobs, byrow = T)
  setup <- setup - min_mat
  lst <- greedy(startup, setup)
  lst$cost <- lst$cost + sum(min_col_val)
  return(lst)
}
```

If we run the function the job sequence becomes:

```{r}
res <- better(startup_costs, setup_costs)
```

That is we use job sequence `r res$seq` with a cost of `r res$cost`.


## Question 2

*Assume that the startup costs are 27, 28, 55, 120 and 26 instead. How will that affect the job sequence using the better strategy?*

We modify the startup costs and run `better`:

```{r}
startup_costs <- c(27, 28, 55, 120, 26)
res <- better(startup_costs, setup_costs)
```

The job sequence now becomes `r res$seq` with a cost of `r res$cost`.


## Question 3

*Make a plot of the job sequence*

```{r, layout="l-body-outset", fig.asp = 0.25, eval=FALSE}
library(ggraph)
library(tidygraph)
edges <- tibble(from = res$seq[1:4], to = res$seq[2:5])
graph <- as_tbl_graph(edges) 
ggraph(graph, layout = 'linear') + 
  geom_edge_link(arrow = arrow(length = unit(4, 'mm')), 
                 end_cap = circle(4, 'mm'), 
                 start_cap  = circle(4, 'mm')) + 
  geom_node_point(colour = "gray", size = 9, alpha = 0.5) +
  geom_node_text(aes(label=name)) +
  theme_graph()
```


```{r Covid opgave 13.5.7}
library(tidyverse)
library(tfa)
dat <- read_csv(system.file("extdata/covid.csv", package = "tfa"))

#Spørgsmål 1: Create a visualization showing the number of cases for each date in Denmark with the following features:

dat %>% 
  filter(country == "Denmark", !is.na(cases)) %>%
  ggplot(aes(x=date, y=cases)) +
  geom_point() +
  geom_line(col="blue")+
  labs(title = "Date and cases compared",
       x = "date",
       y = "cases"
    )


#Spørgsmål 2: Create a visualization showing the total number of deaths per 100000 capita as a function of date with the following features:

dat %>%
   group_by(country) %>%
   mutate(total_deaths = cumsum(replace_na(deaths, 0))) %>%
   mutate(total_deaths_cap = total_deaths/population) %>% 
   ggplot(aes(x = date, y = total_deaths_cap, color = country)) +
   geom_line() +
   labs(title = "Total number of deaths per 100000",
       x = "Date",
       y = "Deaths per 100000 capita",
       color = "Country"
    ) +
   theme(legend.position = "bottom") 



#Spørgmsål 3: Create a visualization showing the percentage of persons vaccinated as a function of date with the following features:

dat %>% 
  filter(year=="2021") %>% 
  mutate(vacperccap = vac/population) %>% 
  mutate(fully = fully_vac/population) %>% 
  pivot_longer(cols = c(vacperccap, fully)) %>% 
  ggplot(aes(x = date, y = value, color = name)) +
  #ggplot(aes(x = date, y = fullyvacperccap, color = country)) +
  facet_wrap(~country) +
  geom_line() 



#Spørgsmål 4: Consider Germany. Create a visualization showing the variation in ICU patients with the following features:

dat %>%
   filter(country == "Germany") %>% 
  ggplot(aes(y = icu_patients)) +
  geom_boxplot() +
  facet_grid(month ~ year, scales = "free_y") +
  labs(title = "ICU patients given year and month",
       x = "",
       y = "Patients") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        legend.position = "bottom")


#Spørgsmål 5: Create a visualization showing the total number of deaths per 100000 capita for each country and year with the following features:
dat %>%
   group_by(country, year) %>% 
   summarise(deaths = sum(deaths/(population/100000) , na.rm = TRUE)) %>% 
   ggplot(aes(y = deaths, x = country, fill = factor(year))) +
   geom_col(position = position_dodge()) +
   labs(title = "Total number of deaths",
    x = "",
    y = "Deaths",
    fill = "Year") +
   theme(legend.position = "bottom")

```



## Colophon 
<!-- Always keep this section for reproducibility -->

This report has been created inside [RStudio](http://www.rstudio.com/ide/) using [R Markdown](https://rmarkdown.rstudio.com/). 

The report was built using:

```{r message = FALSE, warning = FALSE, echo = FALSE}
session <- devtools::session_info()
session$platform
```

Along with these packages:

```{r message = FALSE, warning = FALSE, echo = FALSE}
session$packages %>% 
  select(package, loadedversion, date, source) %>% 
  DT::datatable(rownames = FALSE,
                class = 'cell-border stripe',
                filter = list(position = 'top'),
                options = list(pageLength = 5, 
                           autoWidth = FALSE,
                           bInfo = FALSE,
                           paging = TRUE))
```








